Contextuality and the Born rule in fuzzy quantum theories

### 2017-03-25 prop_letter.tex
prop_letter.tex Updating here and there...

In order to model the idea of finite precision measurement, we replace
the infinity precise~$[0,1]$ by a collection of finite number of
intervals~$\mathscr{I}$, and consider quantum IVPMs~$\bar{\mu}:\events\rightarrow\mathscr{I}$.
Based on the intuitive requirement, $\left[0,0\right]$, $\left[1,1\right]\in\mathscr{I}$,
and operations on intervals,\begin{subequations}
\begin{eqnarray}
 &  & \left[1,1\right]-\left[l,r\right]=\left[1-r,1-l\right]\textrm{ and}\\
 &  & [l_{1},r_{1}]+[l_{2},r_{2}]=[l_{1}+l_{2},r_{1}+r_{2}]
\end{eqnarray}
\end{subequations}a quantum IVPM~$\bar{\mu}$ should satisfy corresponding
axioms:
\begin{itemize}
\item $\bar{\mu}(\mathbb{0})=\left[0,0\right]$. 
\item $\bar{\mu}(\mathbb{1})=\left[1,1\right]$. 
\item For any projection $P$, 
\begin{equation}
\bar{\mu}\left(\mathbb{1}-P\right)=\left[1,1\right]-\bar{\mu}\left(P\right)\textrm{ .}\label{eq:QuantumInterval-valuedProbability-Complement}
\end{equation}
\item For a set of mutually orthogonal projections $\left\{ P_{i}\right\} _{i=1}^{N}$,
we have 
\begin{equation}
\bar{\mu}\left(\sum_{i=1}^{N}P_{i}\right)\subseteq\sum_{i=1}^{N}\bar{\mu}\left(P_{i}\right)\textrm{ .}\label{eq:QuantumInterval-valuedProbability-Inclusion}
\end{equation}
\end{itemize}
Notice that Eq.~(\ref{eq:QuantumInterval-valuedProbability-Inclusion})
relaxes the equality in Eq.~(\ref{eq:QuantumProbability-Addition})
to inclusion because the equality is usually too strong to hold. For
example, if we consider $\left\{ P_{1},P_{2}\right\} =\left\{ P,\mathbb{1}-P\right\} $,
we have 
\[
\bar{\mu}(\mathbb{1})=\left[1,1\right]\subsetneq\left[1+l-r,1+r-l\right]=\bar{\mu}\left(\mathbb{1}-P\right)+\bar{\mu}\left(P\right)
\]
by Eq.~(\ref{eq:QuantumInterval-valuedProbability-Complement}) when
$r>l$.

Although considering IVPM is new for a quantum theory, people have
already defined classical IVPMs~\citep{JamisonLodwick2004}, which
are essentially the same as quantum IVPMs on a restricted mutually
commuting sub-event space~$\events'\subseteq\events$, where the
sub-event space is defined as follow.

\begin{definition}\label{def:sub-event-space}Given a Hilbert space
$\Hilb$, $\events'\subseteq\events$ is called a sub-event space
if it satisfies the follow conditions.
\begin{itemize}
\item $\mathbb{0}\in\events'$.
\item $\mathbb{1}\in\events'$.
\item For any projection $P\in\events'$, we have $\mathbb{1}-P\in\events'$.
\item For a set of mutually orthogonal projections $\left\{ P_{i}\right\} _{i=1}^{N}\subseteq\events'$,
we have $\sum_{i=1}^{N}P_{i}\in\events'$.
\end{itemize}
\end{definition}

On the other hand, Gleason's theorem can be expressed as the limiting
case when the collection of intervals are all singletons~$\mathscr{I}_{\infty}=\set{\left[x,x\right]}{x\in\left[0,1\right]}$.
In the next paragraph, we will show the Kochen-Specker theorem~\citep{kochenspecker1967,peres1995quantum,Redhead1987-REDINA}
is another limiting case when the collection of intervals are $\mathscr{I}_{\mathrm{D}}=\left\{ \imposs,\necess\right\} $,
where $\imposs=\left[0,0\right]$ and $\necess=\left[1,1\right]$.

@@

Since observables in equation~(\ref{eq:MerminSquare}) are not projections,
if we want to reformulate the above argument in the language of quantum
IVPMs, we first need to find the common eigenbasis of each row and
column, and consider the projections to each eigenvectors~\citep{Kernaghan1994,peres1995quantum}.
Let $\mathcal{T}_{24}$ denote all eigenstates listed in table~\ref{table:commonBasis}.
The Kochen-Specker theorem could then be proved by showing there is
no function~$v:\events_{24}\rightarrow\left\{ 0,1\right\} $ satisfying
equation~(\ref{eq:Kochen-Specker-Rules}), where 
\begin{equation}
\events_{24}=\set{\sum_{i=1}^{N}\proj{\psi_{i}}}{\begin{gathered}\textrm{mutually orthogonal }\\
\left\{ \ket{\psi_{i}}\right\} _{i=1}^{N}\subseteq\mathcal{T}_{24}
\end{gathered}
}
\end{equation}
is a sub-event space.

Assume there is a function~$v:\events_{24}\rightarrow\left\{ 0,1\right\} $
satisfying equation~(\ref{eq:Kochen-Specker-Rules}). There must
be a quantum IVPM~$\bar{\mu}_{24}:\events_{24}\rightarrow\mathscr{I}_{\mathrm{D}}$
defined by $\bar{\mu}_{24}\left(P\right)=\iota_{\mathrm{D}}\left(v\left(P\right)\right)$,
where 
\begin{equation}
\iota_{\mathrm{D}}\left(x\right)=\begin{cases}
\necess & \text{ if}x=1\text{ ;}\\
\imposs & \text{ if}x=0\text{ .}
\end{cases}\label{eq:homomorphism-0}
\end{equation}
Hence, there exists an interval-valued frame function~$\bar{f}:\mathcal{T}_{24}\rightarrow\mathscr{I}_{\mathrm{D}}$
defined by $\bar{f}\left(\ket{\psi}\right)=\bar{\mu}_{24}\left(\proj{\psi}\right)$
satisfying the following conditions because of Eq.~(\ref{eq:QuantumInterval-valuedProbability-Inclusion})\begin{subequations}\label{eq:Kernaghan}
\begin{eqnarray}
\necess & \subseteq & \bar{f}(1,0,0,0)+\bar{f}(0,1,0,0)\nonumber \\*
 &  & +\bar{f}(0,0,1,0)+\bar{f}(0,0,0,1)\textrm{ ,}\\
\necess & \subseteq & \bar{f}(1,0,0,0)+\bar{f}(0,1,0,0)\nonumber \\
 &  & +\bar{f}(0,0,1,1)+\bar{f}(0,0,1,-1)\textrm{ ,}\\
\necess & \subseteq & \bar{f}(1,0,0,0)+\bar{f}(0,0,1,0)\nonumber \\*
 &  & +\bar{f}(0,1,0,1)+\bar{f}(0,1,0,-1)\textrm{ ,}\\
\necess & \subseteq & \bar{f}(1,0,0,0)+\bar{f}(0,0,0,1)\nonumber \\*
 &  & +\bar{f}(0,1,1,0)+\bar{f}(0,1,-1,0)\textrm{ ,}\\
\necess & \subseteq & \bar{f}(-1,1,1,1)+\bar{f}(1,-1,1,1)\nonumber \\
 &  & +\bar{f}(1,1,-1,1)+\bar{f}(1,1,1,-1)\textrm{ ,}\\
\necess & \subseteq & \bar{f}(-1,1,1,1)+\bar{f}(1,1,-1,1)\nonumber \\*
 &  & +\bar{f}(1,0,1,0)+\bar{f}(0,1,0,-1)\textrm{ ,}\\
\necess & \subseteq & \bar{f}(-1,1,1,1)+\bar{f}(1,1,-1,1)\nonumber \\*
 &  & +\bar{f}(0,1,1,0)+\bar{f}(1,0,0,-1)\textrm{ ,}\\
\necess & \subseteq & \bar{f}(1,1,-1,1)+\bar{f}(1,1,1,-1)\nonumber \\*
 &  & +\bar{f}(0,0,1,1)+\bar{f}(1,-1,0,0)\textrm{ ,}\\
\necess & \subseteq & \bar{f}(0,1,-1,0)+\bar{f}(1,0,0,-1)\nonumber \\*
 &  & +\bar{f}(1,1,1,1)+\bar{f}(1,-1,-1,1)\textrm{ ,}\\
\necess & \subseteq & \bar{f}(0,0,1,-1)+\bar{f}(1,-1,0,0)\nonumber \\*
 &  & +\bar{f}(1,1,1,1)+\bar{f}(1,1,-1,-1)\textrm{ ,}\\
\necess & \subseteq & \bar{f}(1,0,1,0)+\bar{f}(0,1,0,1)\nonumber \\*
 &  & +\bar{f}(1,1,-1,-1)+\bar{f}(1,-1,-1,1)\textrm{ .}
\end{eqnarray}
\end{subequations}If we sum up equation~\ref{eq:Kernaghan}, we
will have
\begin{eqnarray}
\left[11,11\right] & \subseteq & 4\bar{f}(1,0,0,0)+4\bar{f}(-1,1,1,1)\nonumber \\
 &  & +2\sum_{\ket{\psi}\in\mathcal{T}_{18}}\bar{f}\left(\ket{\psi}\right)\textrm{ ,}\label{eq:QuantumInterval-valuedProbability-Total}
\end{eqnarray}
where 
\begin{eqnarray}
\mathcal{T}_{18} & = & \mathcal{T}_{24}\backslash\{(1,0,0,0),(-1,1,1,1),(1,-1,1,-1),\nonumber \\
 &  & (1,0,-1,0),(1,0,0,1),(1,1,0,0)\}\textrm{ .}
\end{eqnarray}
However, Eq.~(\ref{eq:QuantumInterval-valuedProbability-Total})
can never be satisfied. Because each term in left-hand side is either
$\imposs=\left[0,0\right]$ or $\necess=\left[1,1\right]$, the left-hand
side is a singleton set of an even integer while the right-hand side
is $\left\{ 11\right\} $. Therefore, there is no quantum IVPM~$\bar{\mu}_{24}:\events_{24}\rightarrow\mathscr{I}_{\mathrm{D}}$.
Moreover, there is no quantum IVPM defined on the set of all projectors
as well; otherwise, if we have a quantum IVPM~$\bar{\mu}:\events\rightarrow\mathscr{I}_{\mathrm{D}}$,
then we can restrict $\bar{\mu}$ and define $\bar{\mu}_{24}=\left.\bar{\mu}\right|_{\events_{24}}$.

>>>>

In order to model the idea of finite precision measurement, we replace
the infinity precise~$[0,1]$ by a collection of finite number of
intervals~$\mathscr{I}$, and consider quantum IVPMs~$\bar{\mu}:\events\rightarrow\mathscr{I}$.
Based on the intuitive requirement, $\left[0,0\right]$, $\left[1,1\right]\in\mathscr{I}$,
and operations on intervals,\begin{subequations}\label{eq:interval-operations}
\begin{eqnarray}
 &  & \left[1,1\right]-\left[l,r\right]=\left[1-r,1-l\right]\textrm{ and}\\
 &  & [l_{1},r_{1}]+[l_{2},r_{2}]=[l_{1}+l_{2},r_{1}+r_{2}]
\end{eqnarray}
\end{subequations}a quantum IVPM~$\bar{\mu}$ should satisfy corresponding
axioms:
\begin{itemize}
\item $\bar{\mu}(\mathbb{0})=\left[0,0\right]$. 
\item $\bar{\mu}(\mathbb{1})=\left[1,1\right]$. 
\item For any projection $P$, 
\begin{equation}
\bar{\mu}\left(\mathbb{1}-P\right)=\left[1,1\right]-\bar{\mu}\left(P\right)\textrm{ .}\label{eq:QuantumInterval-valuedProbability-Complement}
\end{equation}
\item For a set of mutually orthogonal projections $\left\{ P_{i}\right\} _{i=1}^{N}$,
we have 
\begin{equation}
\bar{\mu}\left(\sum_{i=1}^{N}P_{i}\right)\subseteq\sum_{i=1}^{N}\bar{\mu}\left(P_{i}\right)\textrm{ .}\label{eq:QuantumInterval-valuedProbability-Inclusion}
\end{equation}
\end{itemize}
Notice that Eq.~(\ref{eq:QuantumInterval-valuedProbability-Inclusion})
relaxes the equality in Eq.~(\ref{eq:QuantumProbability-Addition})
to inclusion because the equality is usually too strong to hold. For
example, if we consider $\left\{ P_{1},P_{2}\right\} =\left\{ P,\mathbb{1}-P\right\} $
and $\bar{\mu}\left(P\right)=\left[l,r\right]$, by Eq.~(\ref{eq:QuantumInterval-valuedProbability-Complement})
and (\ref{eq:interval-operations}), we have 
\[
\bar{\mu}(\mathbb{1})=\left[1,1\right]\subsetneq\left[1+l-r,1+r-l\right]=\bar{\mu}\left(\mathbb{1}-P\right)+\bar{\mu}\left(P\right)
\]
when $r>l$.

Although considering IVPM is new for a quantum theory, people have
already defined classical IVPMs~\citep{JamisonLodwick2004}, which
are essentially the same as quantum IVPMs on a restricted mutually
commuting sub-event space~$\events'\subseteq\events$, where any
subset of $\events$ is called a sub-event space if it satisfies the
follow conditions.
\begin{itemize}
\item $\mathbb{0}\in\events'$.
\item $\mathbb{1}\in\events'$.
\item For any projection $P\in\events'$, we have $\mathbb{1}-P\in\events'$.
\item For a set of mutually orthogonal projections $\left\{ P_{i}\right\} _{i=1}^{N}\subseteq\events'$,
we have $\sum_{i=1}^{N}P_{i}\in\events'$.
\end{itemize}
Quantum IVPM not only stems from classical IVPM but also generalizes
quantum real-valued probability measures which can be expressed a
limiting case when the collection of intervals are all singletons~$\mathscr{I}_{\infty}=\set{\left[x,x\right]}{x\in\left[0,1\right]}$,
where Gleason's theorem is applied. However, it certainly contains
richer phenomena. For example, we will show the Kochen-Specker theorem~\citep{kochenspecker1967,peres1995quantum,Redhead1987-REDINA}
is another limiting case when the collection of intervals are $\mathscr{I}_{\mathrm{D}}=\left\{ \imposs,\necess\right\} $,
where $\imposs=\left[0,0\right]$ and $\necess=\left[1,1\right]$,
in the next paragraph.

@@

Since observables in equation~(\ref{eq:MerminSquare}) are not projections,
if we want to reformulate the above argument in the language of quantum
IVPMs, we first need to find the common eigenbasis of each row and
column, and consider the projections to each eigenvectors~\citep{Kernaghan1994,peres1995quantum}.
Let $\mathcal{T}_{24}$ denote all eigenstates listed in Table~\ref{table:commonBasis}.
The Kochen-Specker theorem could then be proved by showing there is
no function~$v:\events_{24}\rightarrow\left\{ 0,1\right\} $ satisfying
equation~(\ref{eq:Kochen-Specker-Rules}), where 
\begin{equation}
\events_{24}=\set{\sum_{i=1}^{N}\proj{\psi_{i}}}{\begin{gathered}\textrm{mutually orthogonal }\\
\left\{ \ket{\psi_{i}}\right\} _{i=1}^{N}\subseteq\mathcal{T}_{24}
\end{gathered}
}
\end{equation}
is a sub-event space.

Assume there is a function~$v:\events_{24}\rightarrow\left\{ 0,1\right\} $
satisfying equation~(\ref{eq:Kochen-Specker-Rules}). There must
be a quantum IVPM~$\bar{\mu}_{24}:\events_{24}\rightarrow\mathscr{I}_{\mathrm{D}}$
defined by $\bar{\mu}_{24}\left(P\right)=\iota_{\mathrm{D}}\left(v\left(P\right)\right)$,
where 
\begin{equation}
\iota_{\mathrm{D}}\left(x\right)=\begin{cases}
\necess & \text{, if }x=1\text{ ;}\\
\imposs & \text{, if }x=0\text{ .}
\end{cases}\label{eq:homomorphism-0}
\end{equation}
Hence, there exists an interval-valued frame function~$\bar{f}:\mathcal{T}_{24}\rightarrow\mathscr{I}_{\mathrm{D}}$
defined by $\bar{f}\left(\ket{\psi}\right)=\bar{\mu}_{24}\left(\proj{\psi}\right)$
satisfying the following conditions because the right-hand side of
each equation forms an orthogonal basis and we can apply Eq.~(\ref{eq:QuantumInterval-valuedProbability-Inclusion})\begin{subequations}\label{eq:Kernaghan}
\begin{eqnarray}
\necess & \subseteq & \bar{f}(1,0,0,0)+\bar{f}(0,1,0,0)\nonumber \\*
 &  & +\bar{f}(0,0,1,0)+\bar{f}(0,0,0,1)\textrm{ ,}\\
\necess & \subseteq & \bar{f}(1,0,0,0)+\bar{f}(0,1,0,0)\nonumber \\
 &  & +\bar{f}(0,0,1,1)+\bar{f}(0,0,1,-1)\textrm{ ,}\\
\necess & \subseteq & \bar{f}(1,0,0,0)+\bar{f}(0,0,1,0)\nonumber \\*
 &  & +\bar{f}(0,1,0,1)+\bar{f}(0,1,0,-1)\textrm{ ,}\\
\necess & \subseteq & \bar{f}(1,0,0,0)+\bar{f}(0,0,0,1)\nonumber \\*
 &  & +\bar{f}(0,1,1,0)+\bar{f}(0,1,-1,0)\textrm{ ,}\\
\necess & \subseteq & \bar{f}(-1,1,1,1)+\bar{f}(1,-1,1,1)\nonumber \\
 &  & +\bar{f}(1,1,-1,1)+\bar{f}(1,1,1,-1)\textrm{ ,}\\
\necess & \subseteq & \bar{f}(-1,1,1,1)+\bar{f}(1,1,-1,1)\nonumber \\*
 &  & +\bar{f}(1,0,1,0)+\bar{f}(0,1,0,-1)\textrm{ ,}\\
\necess & \subseteq & \bar{f}(-1,1,1,1)+\bar{f}(1,1,-1,1)\nonumber \\*
 &  & +\bar{f}(0,1,1,0)+\bar{f}(1,0,0,-1)\textrm{ ,}\\
\necess & \subseteq & \bar{f}(1,1,-1,1)+\bar{f}(1,1,1,-1)\nonumber \\*
 &  & +\bar{f}(0,0,1,1)+\bar{f}(1,-1,0,0)\textrm{ ,}\\
\necess & \subseteq & \bar{f}(0,1,-1,0)+\bar{f}(1,0,0,-1)\nonumber \\*
 &  & +\bar{f}(1,1,1,1)+\bar{f}(1,-1,-1,1)\textrm{ ,}\\
\necess & \subseteq & \bar{f}(0,0,1,-1)+\bar{f}(1,-1,0,0)\nonumber \\*
 &  & +\bar{f}(1,1,1,1)+\bar{f}(1,1,-1,-1)\textrm{ ,}\\
\necess & \subseteq & \bar{f}(1,0,1,0)+\bar{f}(0,1,0,1)\nonumber \\*
 &  & +\bar{f}(1,1,-1,-1)+\bar{f}(1,-1,-1,1)\textrm{ .}
\end{eqnarray}
\end{subequations}If we sum up equation~\ref{eq:Kernaghan}, we
will have
\begin{eqnarray}
\left[11,11\right] & \subseteq & 4\bar{f}(1,0,0,0)+4\bar{f}(-1,1,1,1)\nonumber \\
 &  & +2\sum_{\ket{\psi}\in\mathcal{T}_{18}}\bar{f}\left(\ket{\psi}\right)\textrm{ ,}\label{eq:QuantumInterval-valuedProbability-Total}
\end{eqnarray}
where 
\begin{eqnarray}
\mathcal{T}_{18} & = & \mathcal{T}_{24}\backslash\{(1,0,0,0),(-1,1,1,1),(1,-1,1,-1),\nonumber \\
 &  & (1,0,-1,0),(1,0,0,1),(1,1,0,0)\}\textrm{ .}
\end{eqnarray}
However, Eq.~(\ref{eq:QuantumInterval-valuedProbability-Total})
can never be satisfied. Because each term in left-hand side is either
$\imposs=\left[0,0\right]$ or $\necess=\left[1,1\right]$, the left-hand
side is a singleton set of an even integer while the right-hand side
is $\left\{ 11\right\} $. Therefore, there is no quantum IVPM~$\bar{\mu}_{24}:\events_{24}\rightarrow\mathscr{I}_{\mathrm{D}}$.
Moreover, there is no quantum IVPM defined on the set of all projectors
as well; otherwise, if we have a quantum IVPM~$\bar{\mu}:\events\rightarrow\mathscr{I}_{\mathrm{D}}$,
then we can restrict $\bar{\mu}$ and define $\bar{\mu}_{24}=\left.\bar{\mu}\right|_{\events_{24}}$.
